\chapter{Mathematical tools} \label{app:math_tools}
This appendix gives a small supplement on the mathematical tools needed to understand the Chapter~\ref{chap:stochasticity}.

\section{Stochastic Diffrential Equations}
The aim of this Section is to give a brief introduction on \emph{Stochastic Differential Equations}.
The discussion is not intended to be extremely formal from a mathematical point of view,
but still provide a complete and consistent exposition of the tools used later in the thesis.
For the readers interested in a formal exposition of the concepts expressed below, please refer to \cite{morters2010brownian}.

This discussion is inspired from \cite{smith2018ito}, article that gives an intuitive description of the differences between different types of stochastic integration,
and provides some physical interpretations of the different possible solutions to stochastic differential equations.

\subsection{Stochasic Processes}
\begin{Definition}[Stochasic Processes]
  A \emph{stochastic process} is \(\{X_t, t\in T\}\) is a collection of random variables living in the same probability space.
  The set \(T\) is usually considered as the \emph{time domain} and can be both discrete or continuous.
\end{Definition}

The most fundamental process is the so-called \emph{Wiener process}, introduced hereafter.
\begin{Definition}[Wiener Process]
  A continuous stochastic process \(\{W_t\}_{t\ge0}\) is called \emph{Wiener process} if it satisfy these properties:
  \begin{itemize}
    \item \(W_0 = 0\);
    \item The increment is indipendent from past values:
          \[\forall u\ge0: \quad W_{t+u} - W_t  \text{ is indipendent from } W_s \quad \forall s\in[0,t];\]
    \item The increment is distributed as a normal variable with variance equal to time difference:
          \[W_{t+u}-W_t \sim \gauss{(0,u)};\]
    \item The process is continuous in time.
  \end{itemize}
\end{Definition}
The importance of this process arises from the fact that it can be used for modeling
most of the stochastic effects that come from the microscopic action of many agents
external to the modeled system.
In the physics context, the Wiener process is often referred as \emph{standard Brownian motion};
moreover, the Wiener process it's used to model the integral white noise.
Lastly, the Wiener process can be derived as the continuous limit of a random walk.

\subsection{Integral over a Wiener Process}
Although a definition of integral for a generic stochastic process is possible,
for this work is interesting only in giving a brief overview of the Wiener process,
because it is the only one needed in the stochastic differential equations of interest.

Let's start by recalling one possible definition of \emph{Riemann Integral} 
of a function \(f\colon[0,T]\to\Real\):
\[
  \int_0^T f{(t)}\dif t \coloneqq \lim_{N\to+\infty} \sum_{k=0}^{N-1}f{(\hat{t}_k)}(t_{k+1}-t_k)
  \quad \text{ with } t_k = \frac{k}{N}T,\,\,\, \hat{t}_k \in [t_k, t_{k+1}].
\]
There exact choices of \(\hat{t}_k\) do not matter, since the limit is converging 
to the same value in any case.

We can now try to extend the Riemann definition to compute the integral of over a Wiener
process. Let \(g:\coloneqq\Real\to\Real\) the function to be integrated. \(g{(t)}\) is a generic (well behaved in some mathematical sense)
random variable depending from time; it's possible to have \(g{(t)} = \tilde{g}{(W_t)}\) or even depending from another stochastic process.
For convenience we will use the convention 
\(\hat{t}_k^\lambda \coloneqq \lambda t_k + (1-\lambda) t_{k+1}\) with \(\lambda\in[0,1]\)\footnote{
  By fixing a value of \(\lambda\) we are essentially choosing \(\hat{t}_k\) for every 
  interval in the summation. In principle, one can think to use different values of \(\lambda\)
  at each different interval, but as it will be clear in a moment, this choice would be
  really impractical and inconvenient.
}. The definition writes as 
\[
  \int_0^T g{(t)}\dif W_t{[\lambda]} \coloneqq \lim_{N\to+\infty} \sum_{k=0}^{N-1}g{(\hat{t}_k^\lambda)}(W_{t_{k+1}}-W_{t_k}).
\]

The first observation to do is that the result won't be a number, as when 
integrating a function, but instead is a random variable, since is calculated as
(limit of) sum of random variables.
\begin{Example}[Integral of Wiener process]
  Let compute the expected value of the integral of \(g{(t)}=2W_t\):
  \[\begin{split}
    \E{\left[\int_0^T 2W_t \dif W_t{[\lambda]}\right]}
      &= \E\left[\lim_{N\to+\infty} \sum_{k=0}^{N-1}2W_{\hat{t}_k^\lambda}(W_{t_{k+1}}-W_{t_k})\right] \\
      &= \E\Bigg[\lim_{N\to+\infty} \sum_{k=0}^{N-1}\Big(2W_{\hat{t}_k^\lambda}W_{t_{k+1}}-2W_{\hat{t}_k^\lambda}W_{t_k} \\
      & \qquad \quad         +(W^2_{t_{k+1}} - W^2_{t_{k+1}}) + (W^2_{\hat{t}_k^\lambda}-W^2_{\hat{t}_k^\lambda}) + (W^2_{t_k}-W^2_{t_k})\Big)\Bigg] \\
      &= \E\left[\lim_{N\to+\infty} \sum_{k=0}^{N-1}\left(-(W_{t_{k+1}}-W_{\hat{t}_k^\lambda})^2 + (W_{\hat{t}_k^\lambda}-W_{t_{k}})^2 + (W^2_{t_{k+1}} - W^2_{t_{k}}) \right)\right]
  \end{split}\]
  The first two addends in the summation are always negative (positive), thus we can commute
  the expected value and the summation; the third term is a telescopic sum
  \[\begin{split}
    \E{\left[\int_0^T 2W_t \dif W_t{[\lambda]}\right]}
      &= \lim_{N\to+\infty} \sum_{k=0}^{N-1}\E\left[-(W_{t_{k+1}}-W_{\hat{t}_k^\lambda})^2 + (W_{\hat{t}_k^\lambda}-W_{t_{k}}\right] + \E\left[W^2_{T}\right] \\
      &= \lim_{N\to+\infty} \sum_{k=0}^{N-1}\left(-(t_{k+1}-\hat{t}_k^\lambda) + (\hat{t}_k^\lambda-t_{k})^2\right) + T \\
      &= T + \lim_{N\to+\infty} \sum_{k=0}^{N-1}\left(-t_{k+1}+2\lambda t_k + 2(1-\lambda)t_{k+1}-t_k\right) \\
      &= T + \lim_{N\to+\infty} \sum_{k=0}^{N-1}(1-2\lambda)(t_{k+1}-t_k) = 2(1-\lambda)T
  \end{split}\]
\end{Example}
It is evident that the choice of \(\lambda\) it's important to define the result.
One can choose infinitely many different stochastic integrations, but the two most common
are \emph{Stratonovich Integration} and \emph{Itô Integration}. They lead to a different result,
with different general properties. As presented in the following sections, the choice of a specific
definition of integration has to be made based on the properties that the modelization requires.

\subsubsection{Stratonovich Integration}
The first definition of integration is the one obtained for \(\lambda=\frac12\).
\begin{Definition}[Stratonovich Integration]
  The \emph{Stratonovich Integration} of a function \(g{(t)}\) over Wiener process is
  \[\int_0^T g{(t)}\circ\dif W_t \coloneqq \lim_{N\to+\infty} \sum_{k=0}^{N-1}g{\left(\frac{t_{k}+t_{k+1}}{2}\right)}(W_{t_{k+1}}-W_{t_k}).\]
\end{Definition}

\subsubsection{Itô Integration}
The second different definition of integration is obtained with \(\lambda=1\).
\begin{Definition}[Itô Integration]
  The second definition of integration is obtained by choosing \(\lambda=1\).
  The \emph{Itô Integration} of a function \(g{(t)}\) over Wiener process is
  \[\int_0^T g{(t)}\dif W_t \coloneqq \lim_{N\to+\infty} \sum_{k=0}^{N-1}g{(t_k)}(W_{t_{k+1}}-W_{t_k}).\]
\end{Definition}

\subsection{Stochastic Differential Equations}
The \emph{stochastic differential equation} arises from the need to describe the
presence of intrinsic noise that affects evolution. Let \(X\) a quantity whose
time evolution is regulated by a deterministic term \(\mu{(X,t)}\), as well as a noisy
term \(\xi_t\). We would like to write an equation similar to
\[\dod{X}{t} = \mu{(X,t)} \textcolor{red}{+ \sigma{(X,t)}\xi_t},\]
where the noisy term is highlighted because we have not a mathematical definition
of it. Of course, \(\xi_t\) must be some sort of stochastic process as defined in 
the previous section. This lead to the first important conclusion: the solution of the
equation will not be a function, but a stochastic process too.

Let's try to make the difinition more formal and give a definition for \(\xi_t\). 
We require the following properties to be satisfied:
\begin{enumerate}
  \item the noise is zero-mean \[\E\left[\xi_t\right] = 0 \quad \forall t;\]
  \item the noise is completely uncorrelated  \[\E\left[\xi_t\xi_{t'}\right] = 0 \quad \forall t\neq t';\]
  \item the noise is stationary (denoting by \(p\) the probability density function)
        \[p{(\xi_{t_1},\dots,\xi_{t_k})} = p{(\xi_{t_1 + t},\dots,\xi_{t_k + t})} \quad \forall t.\]
\end{enumerate}
The generality is not lost with the first property, since the expected value produces a drift,
that can be included in the function \(\mu\). Also, the third property is not restricting too much 
the type of noise, since \(\sigma\) can incorporate an eventual variation of the variance\footnote{
  Of course, property 3 is not just asking to have stationary variance, but the stationarity
  involves the whole distribution.
}. The second property is saying that the noise has no memory of what happened in the past;
the memoryless noise is of course just an approximation of a real physics system, but it is still
good enough to model many systems. Moreover, the noise considered in this work is artificially 
generated and it fulfills this property, so the general case is not treated here.

As we have already mentioned, \(\xi_t\) is not an easy mathematical object to describe, given its non-continuous properties. We, therefore, prefer to analyze its time-integrated version, which we know is a Wiener process.
The differential equation above can be rewritten in terms of differentials (stochastic and nonstochastic) 
\[\dif X = \mu{(X,t)}\dif t + \sigma{(X,t)}\dif W_t,\]
where \(\dif W_t\) it's a Wiener process of duration \(\dif t\).
This is the standard form of a stochastic differential equation describing a generic Brownian motion. It can be alternately integrated with Itô or Stratonovich, leading to two solutions with different properties.


\subsection{Numerical integration}
The Euler method for first-order numerical integration of ordinary differential equations can be extended to include stochastic differentials.
The core rule doing the substitution
\[
  \dif W_t = \xi \sqrt{\dif t} \quad \text{where } \xi \sim \gauss{(0,1)}.
\]
For reference, see \cite{wiki2022euler}.

\section{Generalized Hypergeometric functions}
We include a small appendix where we describe generalized hypergeometric functions.
These functions are often the result of computing statistics on stochastic processes,
such as first passage times.

\begin{Definition}[Pochhammer symbol]
  The Pochhammer symbol of \(a\in\Real\) is defined as 
  \[
    (a)_n \coloneqq \frac{\Gamma(a+n)}{\Gamma(a)},
  \]
  where \(\Gamma\) is the \emph{gamma function} and \(n\ge0\) is an integer.
\end{Definition}

\begin{Definition}[Generalized Hypergeometric function]
  The generalized hypergeometric function \(\prescript{}{p}F_q{\left(a_1,\dots a_p; b_1,\dots,b_q;z\right)}\) is
  \[
    \prescript{}{p}F_q{\left(a_1,\dots a_p; b_1,\dots,b_q;z\right)} =
    \sum_{n=0}^{+\infty} \frac{(a_1)_n\cdots(a_p)_n}{(b_1)_n\cdots(b_q)_n}\frac{z^n}{n!}.
  \]
\end{Definition}

